---
title: "Financial Literacy and Cryptocurrency"
author: "Kieran Yuen"
font-family: Corbel
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# r setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in NFCS data
```{r Loading in NFCS data, include=FALSE}
library(readxl)
X04092022_NFCS_2018_State_Tracking_Data_190623 <- read_excel("04092022_NFCS 2018 State Tracking Data 190623.xlsx")
```

# Cleaning Data
```{r Cleaning Data, include=FALSE}
library(tidyverse)
NFCS_data <- X04092022_NFCS_2018_State_Tracking_Data_190623 %>% 
  select(TRACK,
         NFCSID,
         STATEQ,
         A3,
         A3Ar_w,
         A5_2015,
         A8,
         A6,
         A11,
         M4,
         M6,
         M7,
         M8,
         M9,
         M10) %>%
  
  rename(Year = TRACK,
         NFCS_ID = NFCSID,
         State = STATEQ,
         Gender = A3,
         Age = A3Ar_w,
         Education = A5_2015,
         Income = A8,
         Marital_Status = A6,
         Children = A11,
         Financial_Education = M4,
         Question_1 = M6,
         Question_2 = M7,
         Question_3 = M8,
         Question_4 = M9,
         Question_5 = M10) %>%
  
  na.omit() %>%
  
  filter(Year =="2018"|
           Year == "2015") %>%
  
  mutate(Gender = recode(Gender,
                      '1' = 1,
                      '2' = 0)) %>%
  # mutate(Age = recode(Age,
  #                     '1' = "18-24",
  #                     '2' = "25-34",
  #                     '3' = "35-44",
  #                     '4' = "45-54",
  #                     '5' = "55-64",
  #                     '6' = "65+")) %>%
  
  mutate(Age_24_or_younger = Age == 1,
         Age_24_or_younger = ifelse(Age_24_or_younger == TRUE, 1, 0)) %>%
  
  mutate(Age_34_or_younger = (Age == 1 | Age == 2),
         Age_34_or_younger = ifelse(Age_34_or_younger == TRUE, 1, 0)) %>%
  
  mutate(Age_44_or_younger = (Age == 1 | Age == 2 | Age == 3),
         Age_44_or_younger = ifelse(Age_44_or_younger == TRUE, 1, 0)) %>%
  
  mutate(Education = recode(Education,
                      '1' = 1,
                      '2' = 2,
                      '3' = 3,
                      '4' = 4,
                      '5' = 5,
                      '6' = 6,
                      '7' = 7,
                      '99' = NA_real_)) %>%
  
  mutate(Income = recode(Income,
                      '1' = 1,
                      '2' = 2,
                      '3' = 3,
                      '4' = 4,
                      '5' = 5,
                      '6' = 6,
                      '7' = 7,
                      '8' = 8,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Marital_Status = recode(Marital_Status,
                      '1' = "Married",
                      '2' = "Single",
                      '3' = "Divorced",
                      '4' = "Divorced",
                      '5' = "Widowed/widower",
                      '99' = 'NA')) %>%
  
  mutate(Children = recode(Children,
                      '1' = 1,
                      '2' = 2,
                      '3' = 3,
                      '4' = 4,
                      '5' = 0,
                      '6' = 0,
                      '99' = NA_real_)) %>%
  
    mutate(Financial_Education = recode(Financial_Education,
                      '1' = 1,
                      '2' = 2,
                      '3' = 3,
                      '4' = 4,
                      '5' = 5,
                      '6' = 6,
                      '7' = 7,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Question_1 = recode(Question_1,
                      '1' = 1,
                      '2' = 0,
                      '3' = 0,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Question_2 = recode(Question_2,
                      '1' = 0,
                      '2' = 0,
                      '3' = 1,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Question_3 = recode(Question_3,
                      '1' = 0,
                      '2' = 1,
                      '3' = 0,
                      '4' = 0,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Question_4 = recode(Question_4,
                      '1' = 1,
                      '2' = 0,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Question_5 = recode(Question_5,
                      '1' = 0,
                      '2' = 1,
                      '98' = NA_real_,
                      '99' = NA_real_)) %>%
  
  mutate(Score = select(., Question_1:Question_5) %>% 
           rowSums(na.rm = TRUE)) %>% 
  
  unite("State_Year", State, Year, remove = FALSE) %>% 
  
  mutate(Crypto_State_Year = ifelse(State_Year == '1_2018'
                                    |State_Year =='10_2018'
                                    |State_Year =='11_2018'
                                    |State_Year =='14_2018'
                                    |State_Year =='30_2015'
                                    |State_Year =='30_2018'
                                    |State_Year =='34_2018'
                                    |State_Year =='39_2018'
                                    |State_Year =='43_2015'
                                    |State_Year =='43_2018'
                                    |State_Year =='45_2018'
                                    |State_Year =='46_2015'
                                    |State_Year =='46_2018'
                                    |State_Year =='48_2018'
                                    |State_Year =='7_2015'
                                    |State_Year =='7_2018'
                                    ,1,0)) 
```



# Changing Data Types for variables
```{r Changing Data Types for variables, include=FALSE}
#Gender
NFCS_data$Gender <- as.factor(NFCS_data$Gender)

#Age
NFCS_data$Age <- as.integer(NFCS_data$Age)

NFCS_data$Age_24_or_younger <- as.factor(NFCS_data$Age_24_or_younger)

NFCS_data$Age_34_or_younger <- as.factor(NFCS_data$Age_34_or_younger)

NFCS_data$Age_44_or_younger <- as.factor(NFCS_data$Age_44_or_younger)

#Education
NFCS_data$Education <- as.integer(NFCS_data$Education)

#Income
NFCS_data$Income <- as.integer(NFCS_data$Income)

#Marital Status
NFCS_data$Marital_Status <- as.factor(NFCS_data$Marital_Status)
NFCS_data$Marital_Status <- factor(NFCS_data$Marital_Status,
                              levels = c("Single",
                                         "Married",
                                         "Divorced",
                                         "Separated",
                                         "Widowed/widower"))

# Children
NFCS_data$Children <- as.integer(NFCS_data$Children)

#Financial Education
# NFCS_data$Financial_Education <- as.numeric(NFCS_data$Financial_Education)

#Score
NFCS_data$Score <- as.integer(NFCS_data$Score)

#Crypto_State_Year
NFCS_data$Crypto_State_Year <- as.factor(NFCS_data$Crypto_State_Year)
```


# Regression Models
## Model #1: Demographic information only
```{r Model #1: Demographic information only}
Model_1 <- lm(Score ~ 
                Gender
              + Age 
              + Income 
              + Education 
              + Marital_Status 
              + Children
              , data = NFCS_data)
summary(Model_1)

plot(Model_1)
```

### Model #1: Multicollinearity Check
```{r Model #1: Multicollinearity Check}
car::vif(Model_1)
#variance inflation factor: measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model
#The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014).
```

## Model #2: Demographic information & Crypto State-Year Variable
```{r Model #2: Demographic information & Crypto State-Year Variable}
Model_2 <- lm(Score ~ Gender
              + Age 
              + Income 
              + Education 
              + Marital_Status 
              + Children 
              + Crypto_State_Year
              , data = NFCS_data)
summary(Model_2)

plot(Model_2)
```


### Model #2: Multicollinearity Check
```{r Model #2: Multicollinearity Check}
car::vif(Model_2)
#variance inflation factor: measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model
#The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014).
```



## Model #3: Demographic information & Crypto State-Year Variable & interaction of Age and Crypto State-Year variable
```{r Model #3: Demographic information & Crypto State-Year Variable & interaction of Age and Crypto State-Year variable}
Model_3 <- lm(Score ~ Gender
              + Age
              + Income 
              + Education 
              + Marital_Status 
              + Children 
              + Crypto_State_Year 
              + Age_24_or_younger*Crypto_State_Year
              , data = NFCS_data)
summary(Model_3)

plot(Model_3)
```

```{r Model #4: Demographic information & Crypto State-Year Variable & interaction of Age and Crypto State-Year variable}
Model_4 <- lm(Score ~ Gender
              + Age
              + Income 
              + Education 
              + Marital_Status 
              + Children 
              + Crypto_State_Year 
              + Age_34_or_younger*Crypto_State_Year
              , data = NFCS_data)
summary(Model_4)

plot(Model_4)
```

```{r Model #5: Demographic information & Crypto State-Year Variable & interaction of Age and Crypto State-Year variable}
Model_5 <- lm(Score ~ Gender
              + Age
              + Income 
              + Education 
              + Marital_Status 
              + Children 
              + Crypto_State_Year 
              + Age_44_or_younger*Crypto_State_Year
              , data = NFCS_data)
summary(Model_5)

plot(Model_5)
```








### Model #3: Multicollinearity Check
```{r Model #3: Multicollinearity Check}
car::vif(Model_3)
#variance inflation factor: measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model
#The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014).
```
